Changing Flights In The Air\\ (Abortable Chain Agreement)

### 2015-05-14 paper/intro.tex
Work on the intro, added /web/

However, devising new SMR algorithms, optimized with particular situations or metrics in mind, is hard: today, more than two decades after the first SMR algorithms \cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new SMR algorithms is still hard enough to justify a publication in a top conference \cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

The problem is that it is hard to reuse existing algorithms and ideas because SMR algorithms are not modular. When devising a new algorithm, all the careful analysis, tests, and proofs have to be done again from scratch.

@@

To demonstrate the practical applicability of our framework we provide composable versions of the generalized variations of popular consensus algorithms: classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated Paxos. We have checked the correctness of these algorithms with the TLC model-checker and we provide all the specifications needed to replicate these experiments.

Our framework is also fully formalized and its composition theorem proved in Isabelle/HOL, a foundational interactive theorem prover, thereby certifying the correctness of the composition theorem at the deepest level.

>>>>

However, devising new SMR algorithms, optimized with
particular situations or metrics in mind, is hard: today,
more than two decades after the first SMR algorithms
\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new
SMR algorithms is still complex enough to justify a publication in a top conference
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

The problem is that it is difficult to reuse existing algorithms and ideas because
SMR algorithms are not modular. When devising a new algorithm, all the careful
analysis, tests, and proofs have to be done again from scratch.

@@

To demonstrate the practical applicability of our framework we provide
composable versions of the generalized variations of popular consensus
algorithms: classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated
Paxos. We have checked the correctness of these algorithms with the TLC
model-checker and we provide all the specifications needed to replicate these
experiments. Our framework is also fully formalized and its composition theorem
proved in Isabelle/HOL, a foundational interactive theorem prover, thereby
certifying the correctness of the composition theorem at the deepest level.

### 2015-05-14 paper/paper.tex
Work on the introduction.
Work on the intro, added /web/

\addbibresource{./managed.bib}

@@

\author{
  Rachid Guerraoui\\
  \texttt{rachid.guerraoui@epfl.ch}
  \and
  Giuliano Losa\\
  \texttt{giuliano.losa@epfl.ch}
}

@@

\vspace{2cm}

\begin{center}
This paper is a regular submission.

\vspace{1cm}
This paper is a student paper.
\end{center}

@@

\input{intro}
\input{roadmap}

@@

\end{document}
    We present a new abstraction, Abortable Generalized Consensus, which enables the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms.

    integrating state of the art techniques to optimize executions.  ACA is the most complete approach to date enabling the
    correct-by-construction composition of heterogeneous State-Machine
    Replication algorithms. We provide two specifications satisfying
    ACA and we illustrate their power by refining them to
    obtain a versatile adaptive algorithm that combines five existing
    algorithms into a single one. ACA is completely formalized and its
    underlying composition theorem has been mechanically checked by an interactive
    proof assistant.
}

\vspace{2cm}

\begin{center}
This paper is a regular submission.

\vspace{1cm}
This paper is a student paper.
\end{center}

\section{Introduction}

The State-Machine Replication technique (abbreviated SMR) allows
building reliable services on top of unreliable hardware. The basic idea is to 
replicate a service over several servers, each executing the same
sequence of (deterministic) commands. 
Making SMR algorithms efficient has been a long-standing
challenge and a multitude of algorithms have been proposed. Examples
include Fast Paxos \cite{Lamport06FastPaxos}, Disk Paxos
\cite{GafniLamport03DiskPaxos}, Chain Replication
\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
Ring Paxos
\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos
\cite{CamargosSchmidtPedone07MulticoordinatedPaxos}, Cheap Paxos
\cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting},
Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
and Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency}.

Each of the existing algorithms is best suited to a particular range
of environments, characterized, e.g., by client load, network
topology, likelihood of faults, communication latency and bandwidth, processing
power of nodes, availability of disk storage and memory.
However, the environment faced by real systems evolves over time:
hardware is upgraded, new data-centers are built, the popularity of a
service changes, etc.  Consistent performance therefore comes with
adaptation to change.  Unfortunately, existing SMR algorithms cannot easily be
stopped and replaced by another algorithm on the fly. Moreover,
devising a new algorithm that can change behavior dynamically is hard:
dynamically switching from one SMR algorithm to another requires
synchronizing notoriously intricate algorithms.

We propose a new abstraction, \emph{Abortable Chain Agreement}
(abbreviated ACA), which subsumes SMR and whose implementations,
called \emph{ACA rounds}, can, by construction, pass the baton from
one to another on the fly and efficiently. Moreover, optimizing a replicated
service consisting of several ACA rounds becomes just a matter of devising a new
ACA round, which can then be plugged-in on the fly: any new ACA round
is compatible with every existing ACA round by construction.

Abortable Chain Agreement is the \emph{most complete framework to date} enabling
correct by construction composition of heterogeneous rounds.  
ACA is completely formalized and its underlying composition theorem has been
mechanically checked by an interactive proof assistant.  
Compared to ACA,
existing approaches
\cite{BarNoyETAL87ShiftingGearsChangingAlgorithmsFlyTo,GuerraouiETAL10Next700BftProtocols,GuerraouiKuncakLosa12SpeculativeLinearizability}
are limited. In particular, they do not allow the disorderly execution of
commuting and read-only commands, an important practical optimization
which is at the root of recent work on SMR like Generalized Consensus
\cite{Lamport05GeneralizeConsensus} and Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}.

\begin{comment}
On top of the traditional roles of proposer, acceptor, and learners
\cite{lamport2001paxos} used in consensus algorithms, ACA introduces
the notion of \emph{switchers} and of \emph{scheduling policy}. The switchers
have the task of aborting a running ACA round and initializing a new
ACA round in a safe and live manner.  The \emph{scheduling policy}
decides when to abort an ACA round and choses the next ACA round to
execute. It must ensure that all switchers agree on the next ACA round
to initialize.
\end{comment}

We observe that, because of the FLP impossibility result
\cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess},
most existing SMR algorithms are structured in a sequence of abortable
rounds which can easily be transformed into ACA rounds. To simplify
this task even further, we propose two specifications, \emph{conservative
ACA rounds} and \emph{fast ACA rounds}, which, as we show, can be concretized
to obtain ACA versions of the rounds of five existing algorithms: 
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos. ACA guarantees that the composition of the five ACA
rounds is correct by construction, forming a versatile adaptive replication
algorithm.

The rest of the paper is structured as follows.
\Cref{sec:related} surveys recent related work. \Cref{sec:ca} presents Chain
Agreement, which embodies the requirements for implementing SMR. \Cref{sec:aca}
presents Abortable Chain Agreement and its optimizations to treat commuting
and read-only commands efficiently. \Cref{sec:rounds} presents our specification
of conservative and fast ACA rounds, which implement ACA. In
\cref{sec:examples}, we show that the rounds of  
Classic Paxos, Fast Paxos, Chain Replication, Ring Paxos,
and Multi-Coordinated Paxos can easily be transformed into ACA implementations
by using our specifications of conservative and fast ACA rounds.
Finally, in \cref{sec:isaproofs}, we provide a complete formalization of ACA
and a mechanical proof of the \emph{composition theorem} (\cref{thm:comp}),
checked by the interactive proof assistant Isabelle/HOL \cite{NipkowPaulsonWenzel02IsabelleHOL}.

\begin{thm}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of ACA rounds, the sequence of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Chain Agreement.
\end{thm}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of ACA rounds, the sequence of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of ACA that has a more
  executable form that presented in this section. Moreover,
  instead of sequences of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of ACA uses the theory of I/O-automata, which
  is also formalized in the Isabelle/HOL development. The proof uses a refinement
  mapping from the composition of two consecutive ACA instances to a single ACA
  instance.
\end{proof}

Each Classic Paxos round has a unique leader among the acceptors which
centrally orders the proposed commands, which are accepted by the
other acceptors in the order determined by the leader.  This protocol
ensures that every local acceptor history is a prefix of the local
history of the leader, therefore guaranteeing that the local acceptor histories
form a chain.

Each acceptor broadcasts each newly accepted history to the learners, which
learn a history $h$ when they receive enough extensions of $h$ from the
acceptors.  When instructed by the scheduling policy to abort, a
switcher sends a stop message to the acceptors, which irrevocably stop
updating their local history and send back to the switcher an
acknowledgement containing their local history. A switcher
determines an abort history once it has received acknowledgments from
an abort quorum of acceptors $R$, using the operator $SafeAborts\left(
R \right)$.
acceptors of the preceeding rounds, whereas in ACA rounds, the

>>>>

\addbibresource{managed.bib}

@@

\author{
  Rachid Guerraoui\\
  \texttt{rachid.guerraoui@epfl.ch}
  \and
  Giuliano Losa\\
  \texttt{giuliano@losa.fr}
}

@@

\vspace{2cm}
\begin{center}
This is a regular paper.
\end{center}

@@

\input{intro}

\input{roadmap}

@@
